Estimating the dynamics of SPARQL query results

### 2019-04-11 kgswc19.tex
update

\paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$ -measure for the predictions made considering the union of different combinations of the sets of features (Q, P, R) and varying the window sizes (3, 5, 9 , 17); for reference, we include a baseline that randomly guesses yes / no. 
We can notice the best results were obtained using the characteristics of R reaching values ​​of 0.83 of f1. The results show that predicting changes in query results is indeed a challenging problem. Looking in further detail, as can be expected, prediction performs best when the \ textsc {r} feature - with knowledge of changes in the historical value of results of the input query - is available. And, the features from sets \textsc {q} and  \textsc {p} - which do not assume the availability of such information - obtain less F1-score: more specifically, the best result using \ textsc {pr} is 0.83 with Linear SVM, window size 17, while the best results without \ textsc {r} is 0.60 with Linear SVM, window size 5.

Furthermore, we show in Figure~\ref{fig:result} that effectively the size of the window influences the quality of the results when considering r, being better as the window increases. However, these results are not affected when we consider only predicates in q or p. Also, we show, on the other hand, that all configurations outperform the random baseline.

@@

	\caption{$F_1$-measure for tested classifiers considering different sets of features (Q, P, R) and varying the the window sizes (3,5,9,17)}

@@

\begin{figure}[t]
	%\begin{minipage}[b]{0.30\linewidth}
	\centering
	\begin{tikzpicture}[thick,font=\scriptsize]
	\begin{axis}[
	%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
	xlabel={F1},
	width=5cm, 
	height=5cm,
	ymin=0,
	xtick={1,2,...,4},
	xticklabels={},
	extra x ticks={1,23},
	legend pos=north west,
	ymajorgrids=true,
	grid style=dashed,
	]

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.596)(2,0.6)(3,0.578)(4,0.554)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.519)(2,0.535)(3,0.528)(4,0.511)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.781)(2,0.795)(3,0.797)(4,0.827)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.593)(2,0.59)(3,0.571)(4,0.56)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.775)(2,0.787)(3,0.795)(4,0.819)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.781)(2,0.797)(3,0.8)(4,0.831)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.774)(2,0.787)(3,0.796)(4,0.815)
	};

@@

	\end{axis}
	\end{tikzpicture}
	%		\caption{Evolution of the number of triples on Wikidata.}
	%		\label{graph:datasize}
	%	\end{minipage}
	\hspace{0.1cm}
	%	\begin{minipage}[b]{0.30\linewidth}
	%		\centering
	\begin{tikzpicture}[thick,font=\scriptsize]
	\begin{axis}[
	%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
	xlabel={Precision},
	width=5cm, 
	height=5cm,
	ymin=0,
	xtick={1,2,...,23},
	xticklabels={},
	extra x ticks={1,23},
	legend pos=north west,
	ymajorgrids=true,
	grid style=dashed,
	]

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.598)(2,0.601)(3,0.601)(4,0.619)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.52)(2,0.535)(3,0.528)(4,0.512)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.796)(2,0.798)(3,0.798)(4,0.828)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.596)(2,0.591)(3,0.589)(4,0.579)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.775)(2,0.787)(3,0.795)(4,0.819)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.782)(2,0.797)(3,0.8)(4,0.831)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.774)(2,0.787)(3,0.796)(4,0.815)
	};

@@

	\end{axis}
	\end{tikzpicture}
	%		\caption{Evolution of the number of triples on Wikidata.}
	%		\label{graph:datasize}
	%	\end{minipage}
	\hspace{0.1cm}
	%	\begin{minipage}[b]{0.30\linewidth}
	%		\centering
	\begin{tikzpicture}[thick,font=\scriptsize]
	\begin{axis}[
	%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
	xlabel={Recall},
	width=5cm, 
	height=5cm,
	ymin=0,
	xtick={1,2,...,23},
	xticklabels={},
	extra x ticks={1,23},
	legend pos=north west,
	ymajorgrids=true,
	grid style=dashed,
	]

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.598)(2,0.601)(3,0.579)(4,0.554)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.52)(2,0.535)(3,0.528)(4,0.512)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.787)(2,0.794)(3,0.798)(4,0.828)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.595)(2,0.591)(3,0.572)(4,0.561)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.775)(2,0.787)(3,0.796)(4,0.819)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.783)(2,0.797)(3,0.8)(4,0.831)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.774)(2,0.788)(3,0.796)(4,0.815)
	};

@@

Finally, to predict the TTL of the results of the query -- outputting the number of versions for which the query results can be expected to remain the same -- we employed techniques of multiple linear regression using the \textsc{sklearn}. The latter allows us to see how the periods of estimation and the convergence analysis will behave. We were able to predict the TTL (in relation to the number of invariant weeks) with an average square error of 0.43 weeks and a variance score of 0.01. These results allow to predict future changes in query results with good accuracy using linear regression techniques.

@@

In terms of future work, gold standards based on datasets other than Wikidata could be pursued. Also, our gold standard based on Wikidata could be extended to consider more versions spanning a longer period of time and/or more queries; as previously mentioned, a very large query dataset for Wikidata was recently published~\cite{MalyshevKGGB18} from which it should be possible to extend the Wikidata gold standard (the challenge being to keep a balance of static and dynamic queries with diverse features). Concerning the prediction tasks themselves, there are a variety of other features and approaches that could be applied; one promising direction may be, for example, to apply a more fine-grained analysis of the query, considering for example estimates of the selectivity of particular triples patterns, and combining this information with statistics on the dynamics of the data and the query operators used; our current gold standard could be used to validate such developments. Finally, it would be interesting to investigate the effectiveness of these techniques in practice, developing caching systems, synchronization schedules, and other applications, based on the proposed predictions.

>>>>

\paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$-score for the predictions made considering the union of different combinations of the sets of features (Q, P, R) and varying the window sizes (3, 5, 9, 17); for reference, we include a baseline that randomly guesses yes/no. 
We can note the best results were obtained using the features from \textsc{r} reaching values ​​of 0.83 of $F_1$. The results show that predicting changes in query results is indeed a challenging problem. Looking in further detail, as can be expected, prediction performs best when the \textsc{r} feature - with knowledge of changes in the historical value of results of the input query - is available. And, the features from sets \textsc{q} and  \textsc{p} - which do not assume the availability of such information - obtain less $F_1$-score: more specifically, the best result was using \textsc{pr} (0.83) with Linear SVM and window size 17, while the best results without \textsc{r} was using \textsc{Q} (0.60) with Linear SVM and window size 5.

Furthermore, we show in Figure~\ref{fig:result} that effectively the size of the window influences the quality of the results when considering \textsc{r}, being better as the window increases. However, these results are not affected when we consider only predicates in \textsc{q} or  \textsc{p}. Also, we show, that all configurations outperform the random baseline.

@@

	\caption{$F_1$-score for tested classifiers considering different sets of features (Q, P, R) and window sizes (3,5,9,17)}

@@

\begin{figure}[t]
	%\begin{minipage}[b]{0.30\linewidth}
	\centering
	\begin{tikzpicture}[thick,font=\scriptsize,scale=0.95]
	\begin{axis}[
	%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
	xlabel={F1},
	width=5cm, 
	height=5cm,
	ymin=0,
	xtick={3,5,9,17},
	xticklabels={},
	extra x ticks={3,5,9,17},
	legend pos=north west,
	ymajorgrids=true,
	grid style=dashed,
	]

	\addplot[
	color=red,
	mark=diamond,
	]
	coordinates {
		(3,0.516)(5,0.505)(9,0.515)(17,0.524)
	};

	\addplot[
	color=brown,
	mark=diamond,
	]
	coordinates {
		(3,0.596)(5,0.6)(9,0.578)(17,0.554)
	};

	\addplot[
	color=green,
	mark=diamond,
	]
	coordinates {
		(3,0.519)(5,0.535)(9,0.528)(17,0.511)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(3,0.781)(5,0.795)(9,0.797)(17,0.827)
	};

	\addplot[
	color=black,
	mark=diamond,
	]
	coordinates {
		(3,0.593)(5,0.59)(9,0.571)(17,0.56)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(3,0.775)(5,0.787)(9,0.795)(17,0.819)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(3,0.781)(5,0.797)(9,0.8)(17,0.831)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(3,0.774)(5,0.787)(9,0.796)(17,0.815)
	};

@@

	\end{axis}
	\end{tikzpicture}
	%		\caption{Evolution of the number of triples on Wikidata.}
	%		\label{graph:datasize}
	%	\end{minipage}
	\hspace{0.1cm}
	%	\begin{minipage}[b]{0.30\linewidth}
	%		\centering
	\begin{tikzpicture}[thick,font=\scriptsize,scale=0.95]
	\begin{axis}[
	%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
	xlabel={Precision},
	width=5cm, 
	height=5cm,
	ymin=0,
	xtick={3,5,9,17},
	xticklabels={},
	extra x ticks={3,5,9,17},
	legend pos=north west,
	ymajorgrids=true,
	grid style=dashed,
	]

	\addplot[
	color=red,
	mark=diamond,
	]
	coordinates {
		(3,0.516)(5,0.505)(9,0.515)(17,0.524)
	};

	\addplot[
	color=brown,
	mark=diamond,
	]
	coordinates {
		(3,0.598)(5,0.601)(9,0.601)(17,0.619)
	};

	\addplot[
	color=green,
	mark=diamond,
	]
	coordinates {
		(3,0.52)(5,0.535)(9,0.528)(17,0.512)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(3,0.796)(5,0.798)(9,0.798)(17,0.828)
	};

	\addplot[
	color=black,
	mark=diamond,
	]
	coordinates {
		(3,0.596)(5,0.591)(9,0.589)(17,0.579)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(3,0.775)(5,0.787)(9,0.795)(17,0.819)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(3,0.782)(5,0.797)(9,0.8)(17,0.831)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(3,0.774)(5,0.787)(3,0.796)(4,0.815)
	};

@@

	\end{axis}
	\end{tikzpicture}
	%		\caption{Evolution of the number of triples on Wikidata.}
	%		\label{graph:datasize}
	%	\end{minipage}
	\hspace{0.1cm}
	%	\begin{minipage}[b]{0.30\linewidth}
	%		\centering
	\begin{tikzpicture}[thick,font=\scriptsize,scale=0.95]
	\begin{axis}[
	%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
	xlabel={Recall},
	width=5cm, 
	height=5cm,
	ymin=0,
	xtick={3,5,9,17},
	xticklabels={},
	extra x ticks={3,5,9,17},
	legend pos=north west,
	ymajorgrids=true,
	grid style=dashed,
	]

	\addplot[
	color=red,
	mark=diamond,
	]
	coordinates {
		(3,0.516)(5,0.505)(9,0.515)(17,0.524)
	};

	\addplot[
	color=brown,
	mark=diamond,
	]
	coordinates {
		(3,0.598)(5,0.601)(9,0.579)(17,0.554)
	};

	\addplot[
	color=green,
	mark=diamond,
	]
	coordinates {
		(3,0.52)(5,0.535)(9,0.528)(17,0.512)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(3,0.787)(5,0.794)(9,0.798)(17,0.828)
	};

	\addplot[
	color=black,
	mark=diamond,
	]
	coordinates {
		(3,0.595)(5,0.591)(9,0.572)(17,0.561)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(3,0.775)(5,0.787)(9,0.796)(17,0.819)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(3,0.783)(5,0.797)(9,0.8)(17,0.831)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(3,0.774)(5,0.788)(9,0.796)(17,0.815)
	};

@@

Finally, to predict the TTL of the results of the query -- outputting the number of versions for which the query results can be expected to remain the same -- we employed techniques of multiple linear regression using the \textsc{sklearn} library. The latter allows us to see how the periods of estimation and the convergence analysis will behave. We were able to predict the TTL (in relation to the number of invariant weeks) with an average square error of 0.43 weeks and a variance score of 0.01. These results allow to predict future changes in query results with good accuracy using linear regression techniques.

@@

In terms of future work, gold standards based on datasets other than Wikidata could be pursued. Also, our gold standard based on Wikidata could be extended to consider more versions spanning a longer period of time and/or more queries; as previously mentioned, a very large query dataset for Wikidata was recently published~\cite{MalyshevKGGB18} from which it should be possible to extend the Wikidata gold standard (the challenge being to keep a balance of static and dynamic queries with diverse features). Concerning the prediction tasks themselves, there are a variety of other features and approaches that could be applied; one promising direction may be, for example, to apply a more fine-grained analysis of the query, considering for example estimates of the selectivity of particular triples patterns, and combining this information with statistics on the dynamics of the data and the query operators used; our current gold standard could be used to validate such developments. Finally, it would be interesting to investigate the effectiveness of these techniques in practice, developing caching systems, synchronization schedules, and other applications, based on the evaluated predictions.
